---
title: "MITx-15.071x-U1-Reciation-Adding_data_to_data_frame"
author: "Rahul Yadav"
date: "April 13, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir =  "C:/Users/pySag/Desktop/MITx15.071x")
```


> #### Outline

In this section we are going to add a new variable to our data frame, that takes the value 1 if the food has higher sodium than average, and 0 if the food has lower sodium than average.

To check if the first food in the dataset has higher amount of sodium compared to the average, we can do the following:

* Dig up the 1st value in the Sodium vector.

``` {r}
# Lets first import the dataset.
USDA <- read.csv("C:/Users/pySag/Desktop/MITx15.071x/Datasets/USDA.csv")
USDA $ Sodium[1] > mean( USDA $ Sodium, na.rm = TRUE)

# How about all the food in the sequence till 50.
USDA $ Sodium[1:50] > mean( USDA $ Sodium, na.rm = TRUE)

```

> #### Inference

--- ---

#### The following observations we came up with:

* #### 1st food --> higher sodium content than average.
* #### 50th food --> lower sodium content than average.

--- ---

#### How about all observations:


```{r}
HighSodium <- USDA $ Sodium > mean( USDA $ Sodium, na.rm = TRUE )

# Let's look at the structure of HighSodiuum

str(HighSodium)

# What up with this values:
# The former vector has values either TRUE or FALSE, hence are called logical.
```

--- ---

#### But we want all observations to be either 1 or 0.
Fortunately, R provide us with another function called **as.numeric()**, which is **a type convertion to numeric vector form any non-numeric vector.**

--- ---

```{r}
HighSodium <- as.numeric( USDA $ Sodium > mean( USDA $ Sodium, na.rm = TRUE))
# Let's now check.
str(HighSodium)
```


> ## Adding a vector to a data frame.

--- ---

#### Now the vector, __HighSodium__ is _not associated_ with the "USDA data frame".
#### __Question__ is how can we add the vector "HighSodium" to our USDA data frame.

#### **To do this, simply add '$' with HighSodium with our existing USDA data frame.**

--- ---

```{r}
USDA $ HighSodium <- as.numeric( USDA $Sodium > mean( USDA $ Sodium, na.rm = TRUE))

#check if HighSodium is added to our USDA data frame.
str(USDA)

#Now add other vectors to the data frame as follows, Protein, Carbohydreates, Fats.
USDA $ HighProtein <- as.numeric( USDA $ Protein > mean( USDA $ Protein, na.rm = TRUE))

USDA $ HighFat <- as.numeric( USDA $ TotalFat > mean( USDA $ TotalFat, na.rm = TRUE))

USDA $ HighCarb <- as.numeric( USDA $ Carbohydrate > mean( USDA $ Carbohydrate, na.rm = TRUE))

#check the other variables are added or not.
str(USDA)

```

--- ---

```{r results='hide'}
# Lets save our new data frame.
# We dont want our output to display, as it will too ugly.
write.csv(USDA, file = "C:/Users/pySag/Desktop/MITx15.071x/Datasets/USDA.csv")
```

--- ---

``` {r}
str(USDA)
```



