---
title: "MITx-15.071x-U1-Recitation"
author: "Rahul Yadav"
date: "April 11, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/pySag/Desktop/MITx15.071x" )
```


> ## About Recitation 1

-- --

#### In the following recitation, we get more familiar with R and review some basic functions to read, visualize & analyze data sets.

> ### Objective

-- --

* #### We will be looking at nutritional data over 7K foods distributed by the US Dept. of Agriculture.

* #### Analyzing USDA Food Database in the following recitation.


> ### Analyzing the USDA Food Database.

--- ---

> **Loading the data set**

-- --

* #### We'll start by reading in our dataset USDA.csv.
* #### It contains all foods in the USDA database in 100-gram amounts.

```{r}
getwd()

# read in the data set.

USDA <- read.csv("C:/Users/pySag/Desktop/MITx15.071x/Datasets/USDA.csv")
```

> **Learning about data**

#### We can use the structure function or **str()** & **summary** to get more info. on our USDA data set.

```{r}
str(USDA)

# summary of the data

summary(USDA)
```

-- --


#### Following is the information we obtained from the USDA data set:
  
* ####   **Max. amount of Cholesterol is 3100 mgs.**
* ####   **Whereas, the mean is only 41.55 mgs.**

#### We also have the information about the number of **non-available** entries.

* ####  For instance, about **1910 foods** have _missing entries_ for their **sugar levels.**

* #### _Maximum amount of sodium_ is **38,758** which is high in repect to a daily amount of 25K.
  
#### **Let's Investigate!**


> #### INVESTIGATION

-- --

#### To check the values of sodium levels in the foods within the data set, we can use the following code:
  
  
```{r}
summary(USDA $ Sodium )
```

**Note:** The measurement is in **milligrams or mgs**

> **Which food has the maximum amount of sodium in it?**

##### To the find which food has the maximum amount od sodium in it, we are going to use the following function, **which.max()** followed by the subset vector _sodium_ in our USDA data set.

```{r tidy = TRUE}
which.max(USDA $ Sodium)

# This will throw back the index of the food variable we are intersested to find.

# A neat trick to find the name of the variable if the data set is huge, is names() method. Simply pass our "data frame" as argument in it.

names(USDA)

# now we can find the variable "Description" that will provide us the name of the food that we are interested in.

USDA $ Description[ 265 ]
```

> #### **Observation**


#### It turns out that **"table salt"** contains the *maximum amount* of **sodium** content, which is a no brainer hence, it might be more interesting to find out **which foods for instance contain more than (say) 10K mgs of sodium?**


#### To do the following:
-- --

#### 1. Create a new data frame called **"HighSodium"**.

#### 2. __Subsetting__ With only the *foods that have sodium content* **exceeding 10K.**

-- --

```{r}

# Creating a new subset data fram called -> "Highsodium".
# Subsetting USDA with sodium as the subset parameter and value > 10K.

# This will contain "Observations" listed as foods containing sodium > 10K.
HighSodium <- subset(USDA, Sodium > 10000)

```

-- --

#### We need to see how many ovservations this data frame has, and this can be done by using function **nrow()**.

* #### nrow computes the no. of rows in the data frame HighSodium.

-- --

```{r}
nrow(HighSodium)

# Since there are not many, we can output there names by looking at there discription vector.
```

#### __Note:__ This time we can't simply output using our "old data frame". 

* #### __Why?__, well _Description vector_ now **is not associated with the USDA data frame**.
* #### But rather with **"HighSodium data frame"!**

```{r}
HighSodium $ Description
```

--- ---

#### One of the food's such is caviar is well know to be top amongst food having high sodium content, But it dosen't seem's to appear in this list. **Let's find out**.

#### **Issues:**

* #### We have to track it down in our desctiption variable.
* #### This could easliy become a tedious task.

#### Hence R provides us with a very useful function called, **match()**, let take it for a spin!

-- --

```{r}
match("CAVIAR", USDA $ Description )

# Now let's use the index to find caviar and how much "mgs of sodium"" it contains  / 100 gms of caviar.

USDA $ Sodium[ 4154 ]

# A neat trick to combine the above two step in a single one is as follows:

USDA  $ Sodium[ match( "CAVIAR", USDA $ Description )]
```


> #### INFERENCE

--- ---

* #### contains 1500 mgs of Sodium / 100 grams of caviar.
    * #### This is very small compared to 10K mgs or 38K mgs.
    
* #### But, is it a fair comparison?
    * #### maybe the best way to find "how big the value is" by comparing it to:
        * #### Mean & Standard Deviation.
        
--- ---
        
```{r}
# R summary function will provide us with the "Mean", but not "Standard Deviation". 
summary(USDA $ Sodium )

# To compute the Standard Deviation, recall that "sd()" computes it.
sd( USDA $ Sodium )

```

--- ---

#### **Interesting!** why does it gives us the result "NA"?

* #### Simply because we forgot to remove the non-available entries before computing our satistical measure.

--- ---

``` {r}
sd( USDA $ Sodium, na.rm = TRUE )
```


> ## Conclusion

--- ---

####**Note:** if we sum the mean and the standard deviation, we obtain around 1,400 milligrams, which is still smaller than the amount of sodium in 100 grams of caviar.


#### ***Well, this means that caviar is pretty rich in sodium
compared to most of the foods in our data set***

--- ---
