---
title: "MITx15.071x-Data-Analysis-Plots&SummaryTables"
author: "Rahul Yadav"
date: "April 12, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/pySag/Desktop/MITx15.071x" )
```
--- ---
> ## Loading our data set

```{r}
WHO <- read.csv("C:/Users/pySag/Desktop/MITx15.071x/Datasets/WHO.csv")
```


 ```{r}
 library(plotly)

 ```

-- --

> ## __Visualization tool: The Histogram__

--- ---

```{r}
plot_ly(x = WHO$CellularSubscribers, type = "histogram") 
```

> ## __Visualization toool: The Boxplot__

--- ---

* __A box plot is useful for understanding the statistical range of a variable.__
    * This box plot shows how life expectancy in countries varies according to the region the country is in. 
    * The box for each region shows the _range_ between the _first_ and _third quartiles_ with __the middle line__ marking the __median value__. 

* The __dashed lines__ at the _top and bottom of the box_, often called __whiskers__,
    * show the _range_ from the __minimum to maximum values__,
    * excluding any outliers, which are plotted as circles.
    
* __Outliers__ are _defined_ by first computing
    * the difference between the first and third quartiles, or the height of the box.


```{r}
# Following is the R code to draw a boxplot of Region vs LifeExpectancy.
plot_ly( x = WHO$Region, y = WHO$LifeExpectancy, type = "box", ylab = "Life Expectancy")
```

> ### Understanding our boxplot for Region vs LifeExpectancy

--- ---

This box plot shows us the following:

1. _Europe_ has the __highest median life expectancy__.

2. The _Americas_ has the __smallest inter-quartile range__, and 

3. The _Eastern Mediterranean_ region has the __highest overall range of life expectancy__ values.

-- --

> ## __Summary Tables__

This is similar to what we saw in the summary output and counts the number of observations in each category of Region.

--- --- 

```{r}
table(WHO $ Region)
```
    
__Note:__ Tables work well for variables with only a few possible values.

-- --

> ### Extracting _more information_ on __Numerical Variables__ using __tapply()__.

``` {r}
## syntax: tapply( vector, step by, function )
tapply(WHO$Over60, WHO$Region, mean)
```

__Note/FIX__: If somehow tapply throw back an error of "length not matching in arguments, just import your data set through r console. Make sure it's in the global environment.

-- --

> **Understanding tapply()**

In general, **tapply() splits the data by the second argument, then applies the third argument function, to the variable given as the first argument** as formerly mentioned in the chunk above.

Here's the following explaination of the result we obtained from the above code chunk we executed:

1. The mean % of population in African region is --> 5%
2. The mean % of population "over 60" in European countries is --> 20%.

To understand better, here's another example:

``` {r}
tapply( WHO $ LiteracyRate, WHO $ Region, min)
```

--- --- 

> ## Handling missing values in our data set.

If we closely go throught the output churned by the above chunk, there's **data marked as NA(missing)**, this is not good for data analysis, why? well here are the following reasons:

1."missing data" may give us __incorerct results__.
2. worse "mislead" us into persuing something undesirable. 

####Question is how do we take care of it?

Well there are techniques through which R allows us to handle such issues easily.

```{r}
tapply(WHO $ LiteracyRate, WHO $ Region, min, na.rm = TRUE )
```


> **Quiz** 

Use the tapply function to find the average child mortality rate of countries in each region.

Which region has the lowest average child mortality rate across all countries in that region?

```{r}
tapply(WHO $ ChildMortality, WHO $ Region, mean )
```